---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2023-05-10"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)

library(R.utils)

library(pROC)

library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)

library(mvtnorm)
```

```{r setup, include = FALSE} 
plotFolder <- here("images")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)
```


```{r}
R.files.sources <- c(list.files("../R", pattern="*.R$", full.names=TRUE),
                     list.files("../simu_R", pattern="*.R$", full.names=TRUE))
print(R.files.sources)

sapply(R.files.sources, source)
```

# Simulation 1
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-5 + W + 2.25A -0.5WA)]$

Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys1}
simu.num = 1
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```

# Simulation 2
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1.3^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2.5 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-7 + 3W + 5sin(1.25A^{1.5}) + 5A + 3WA)]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys2}
simu.num = 2
```


```{r child = 'Simu_results_template_one_simu.Rmd'}
```

# Simulation 3
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-10 - 3W + 4A + \mathbf{I}(A>2) * 5sin((0.8A)^2 - 2.6)  )]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys3}

generate_data_3 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- as.numeric(U_Y < plogis(-10 - 3*W + 4*A + 5*sin((0.8*A)^2-2.6)*as.numeric(A > 2)) )
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


DGS <- generate_data_3
sn = 3
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```


# Simulation 4
```{r}
write_matex2 <- function(x) {
  begin <- "\\begin{bmatrix}"
  end <- "\\end{bmatrix}"
  X <-
    apply(x, 1, function(x) {
      paste(
        paste(x, collapse = "&"),
        "\\\\"
      )
    })
  paste(c(begin, X, end), collapse = "")
}

U_W_var <- matrix(c(1,        0.8,      0.9*1.5,     0.2*0.5,      0.1*0.3,
                      0.8,      1,        0.8*1.5,     0.3*0.5,      0.1*0.3, 
                      0.9*1.5,  0.8*1.5,  1.5^2,       0.1*0.5*1.5,  0.5*0.3*1.5,
                      0.2*0.5,  0.3*0.5,  0.1*0.5*1.5, 0.5^2,        0.75*0.3*0.5,
                      0.1*0.3,  0.1*0.3,  0.5*0.3*1.5, 0.75*0.3*0.5, 0.3^2), nrow = 5)

U_W_mu <- matrix(c(0, 0, 5, 1, 1), nrow=5)
```

Data structure:  $O = (W_1, W_2, W_3, W_4, W_5, A, Y)$

 * W - baseline covariates
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$

* Structural equations F and endogenous variables:
  + $W \sim  N(\mu_W, \Sigma_W)$
  + $A = bound(0.1W_1 + 0.2W_2 + 0.5W_3 + 0.15W_4 - 0.05W_5 - 0.01W_3W_5 + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-7 -W_1 + 2W_2 - 0.5W_4 - W_1W_3 + 4A + 0.5AW_2)]$
    - where $\mu_W = `r write_matex2(U_W_mu)`$, and $\Sigma_W = `r write_matex2(U_W_var)`$

Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys4}

generate_data_4 <- function(n, a=NA){
  # exogenous variables
  U_W_var <- matrix(c(1,        0.8,      0.9*1.5,     0.2*0.5,      0.1*0.3,
                      0.8,      1,        0.8*1.5,     0.3*0.5,      0.1*0.3, 
                      0.9*1.5,  0.8*1.5,  1.5^2,       0.1*0.5*1.5,  0.5*0.3*1.5,
                      0.2*0.5,  0.3*0.5,  0.1*0.5*1.5, 0.5^2,        0.75*0.3*0.5,
                      0.1*0.3,  0.1*0.3,  0.5*0.3*1.5, 0.75*0.3*0.5, 0.3^2), nrow = 5)
  U_W <- rmvnorm(n, mean = c(0, 0, 5, 1, 1), sigma = U_W_var)
  U_A <- rnorm(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- data.frame(U_W)
  colnames(W) <- paste0('W', 1:ncol(W))
  
  if(is.na(a)){
    A <-  (0.1*W$W1 + 0.2*W$W2  + 0.5*W$W3 + 0.15*W$W4 - 0.05*W$W5 - 0.01*W$W3*W$W5 + U_A)
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }

  Y <- as.numeric(U_Y < plogis(-7 - W$W1 + 2*W$W2 - 1*W$W1 * W$W3 - 0.5*W$W4 + 4*A + 0.5*A*W$W2 ))
  mean(Y)
  
  # data frame
  O <- W
  O$A = A
  O$Y = Y
  return(O)
}


DGS <- generate_data_4
sn = 4
```

# ```{r child = '2_1_Simu_results_template_0.Rmd'}
# ```

# Simulation 5

Data structure:  $O = (W_1, W_2, W_3, W_4, W_5, A, Y)$

 * W - baseline covariates
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$

* Structural equations F and endogenous variables:
  + $W \sim  N(\mu_W, \Sigma_W)$
  + $A = bound(0.1W_1 + 0.2W_2 + 0.5W_3 + 0.15W_4 - 0.05W_5 - 0.01W_3W_5 + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-10 -W_1 + 2W_2 - 0.5W_4 - 0.5W_1W_3 + 4A + \mathbf{I}(A>2) *5sin((0.8A)^2 - 2.6))]$
    - where $\mu_W = `r write_matex2(U_W_mu)`$, and $\Sigma_W = `r write_matex2(U_W_var)`$

Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys5}
generate_data_5 <- function(n, a=NA){
  # exogenous variables
  U_W_var <- matrix(c(1,        0.8,      0.9*1.5,     0.2*0.5,      0.1*0.3,
                      0.8,      1,        0.8*1.5,     0.3*0.5,      0.1*0.3, 
                      0.9*1.5,  0.8*1.5,  1.5^2,       0.1*0.5*1.5,  0.5*0.3*1.5,
                      0.2*0.5,  0.3*0.5,  0.1*0.5*1.5, 0.5^2,        0.75*0.3*0.5,
                      0.1*0.3,  0.1*0.3,  0.5*0.3*1.5, 0.75*0.3*0.5, 0.3^2), nrow = 5)
  U_W <- rmvnorm(n, mean = c(0, 0, 5, 1, 1), sigma = U_W_var)
  U_A <- rnorm(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- data.frame(U_W)
  colnames(W) <- paste0('W', 1:ncol(W))
  
  if(is.na(a)){
    A <-  (0.1*W$W1 + 0.2*W$W2  + 0.5*W$W3 + 0.15*W$W4 - 0.05*W$W5 - 0.01*W$W3*W$W5 + U_A)
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }

  Y <- as.numeric(U_Y < plogis(-10 - W$W1 + 2*W$W2 - 0.5*W$W1 * W$W3 - 0.5*W$W4 + 4*A + 5*sin((0.8*A)^2-2.6)*as.numeric(A > 2)) )
  mean(Y)
  
  # data frame
  O <- W
  O$A = A
  O$Y = Y
  return(O)
}

DGS <- generate_data_5
sn = 5
```

# ```{r child = '2_1_Simu_results_template_0.Rmd'}
# ```

# Simulation 6

Data structure:  $O = (W_1, W_2, W_3, A, Y)$

 * W - baseline covariates
 * A - treatment level based on W, continuous between 0 and 1
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Structural equations F and endogenous variables:
  + $W_1 \sim Uniform(o,1)$
  + $W_2 \sim Bernoulli(\mu=0, \sigma^2 = 2^2)$
  + $W_3 \sim  N(W_1, 0.25*exp(2W_1))$
  + $A \sim Beta(v(W)\mu(W), v(W)[1-\mu(W)])$
  + $Y \sim Bernoulli(Q_0(A,W))$
    - where:
    - $v(W) = exp(1 + 2W_1expit(W3))$
    - $\mu(W) = expit(0.03 - 0.8log(1+W_2) + 0.9exp(W_1)W_2 - 0.4arctan(W_3+2)W_2W_1)$
    - $\bar{Q}_0(A,W) = expit(-2 + 1.5A + 5A^3 - 2.5W_1 + 0.5AW_2 - log(A)W_1W_2 + 0.5A^{3/4}W_1W_3)$

Outcome of interest: $E_0[Y|a,W]$, $a \in (0,1])$, the causal dose-response curve

```{r check_sys6}
expit <- function(x){
  return(exp(x)/(1+exp(x)))
}

generate_data_6 <- function(n, a=NA){
  W1 <- runif(n)
  W2 <- rbinom(n, size=1, prob =0.7)
  W3 <- rnorm(W1, 0.25*exp(2*W1))
  
  v_W <- exp(1 + 2*W1*expit(W3))
  mu_W <- expit(0.03 - 0.8*log(1+W2) + 0.9*exp(W1)*W2 - 0.4*atan(W3+2)*W2*W1)
  if(is.na(a)){
    A <- rbeta(n, v_W*mu_W, v_W*(1-mu_W))
  } else {
    A <- a
  }
  

  Q <- expit(-2 + 1.5*A + 5*A^3 - 2.5*W1 + 0.5*A*W2 - log(A)*W1*W2 + 0.5*(A^(3/4))*W1*W3)
  Y <- rbinom(n, size = 1, prob = Q)

  O <- data.frame(W1, W2,W3, A, Y)
  return(O)
}

DGS <- generate_data_6
sn = 6
```

# ```{r child = '2_1_Simu_results_template_s6.Rmd'}
# ```




# Simulation 7

Data structure:  $O = (W, A, Y)$

 * W - baseline covariates
 * A - treatment level based on W, continuous between 0 and 1
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2.5 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-6 + W + 3.5A\mathbf{I}(A\geq2) - 4A\mathbf{I}(A\geq4) - 0.5WA)]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve


```{r check_sys7}
generate_data_7 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 1)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2.5 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- as.numeric(U_Y < plogis(-6 + W + 3.5*A*as.numeric(A >= 2) - 4*A*as.numeric(A >= 4) - 0.5 * W * A ))
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}

DGS <- generate_data_7
sn = 7
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```

# plots
```{r fig.height=6, fig.width=12}
DGS <- generate_data_2
sn = 2
nn= 500
load(here("data", "rdata", paste0("02_simu_V5_sys", sn, "_", nn, ".RData")))

if(sn %in% c(1,7)){
  load(file=here("data", "rdata", paste0("02_simu_V5_sys", sn ,"_psi0.RData")))
  results = run_simu_rep_summary(results)
}

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_2 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_2 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())

legend <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "legend") 




DGS <- generate_data_3
sn = 3
nn=5000
load(here("data", "rdata", paste0("02_simu_V5_sys", sn, "_", nn, ".RData")))

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_3 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_3 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())




DGS <- generate_data_7
sn = 7
nn=500
load(here("data", "rdata", paste0("02_simu_V5_sys", sn, "_", nn, ".RData")))


if(sn %in% c(1,7)){
  load(file=here("data", "rdata", paste0("02_simu_V5_sys", sn ,"_psi0.RData")))
  results = run_simu_rep_summary(results)
}

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_7 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_7 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())



p1 = arrangeGrob(p_est_avg_2, p_est_avg_3, p_est_avg_7, nrow=1, left= "Estimation & CI")
p2 = arrangeGrob(p_cr_2, p_cr_3, p_cr_7, nrow=1, left= "95% CI Coverage Rate")


p <- grid.arrange(p1, p2, legend, 
                  layout_matrix = rbind(c(1,3),
                                        c(2,3)),
                  widths = c(3,0.4),
                  top = "         Simulation 2                                                       Simulation 3                                                        Simulation 4                            \n        n = 500                                                               n = 5000                                                              n = 500                               ")
 

save_plot = here(getwd(), "plots", paste0("simu_sys234_performences_est_cr.png"))
 
ggsave(save_plot, plot=p, width = 12, height = 5, dpi = 800)
```



# colinearity of basis functions
```{r}
library(corrplot)

set.seed(123)

n = 1000
obs = generate_data_1(n)

y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)

hal_CV <- fit_hal(X = X, Y = Y, family = "binomial",
                    return_x_basis = TRUE,
                    num_knots = hal9001:::num_knots_generator(
                      max_degree = ifelse(ncol(X) >= 20, 2, 3),
                      smoothness_orders = 1,
                      base_num_knots_0 = 20,
                      base_num_knots_1 = 20 # max(100, ceiling(sqrt(n)))
                    ))

summary(hal_CV)

x_basis_CV = cbind(1, hal_CV$x_basis)[, which(hal_CV$coefs != 0)]

M_CV <- cor(as.matrix(x_basis_CV))
corrplot(M_CV, method="circle")
# undersmoothing
```

```{r}
hal_undersmooth <- undersmooth_hal(X, Y, fit_init = hal_CV, family = "binomial")
lambda_u_g = hal_undersmooth$lambda_under

if(!is.na(lambda_u_g)){
  hal_u_g <- fit_hal(X = X, Y = Y, family = "binomial",
                         return_x_basis = TRUE,
                         num_knots = hal9001:::num_knots_generator(
                           max_degree = ifelse(ncol(X) >= 20, 2, 3),
                           smoothness_orders = 1,
                           base_num_knots_0 = 20, #200
                           base_num_knots_1 = 20 # max(100, ceiling(sqrt(n)))
                         ),
                         fit_control = list(
                           cv_select = FALSE,
                           n_folds = 10,
                           foldid = NULL,
                           use_min = TRUE,
                           lambda.min.ratio = 1e-4,
                           prediction_bounds = "default"
                         ),
                         lambda = lambda_u_g)
  
  summary(hal_u_g)
  x_basis_U = cbind(1, hal_u_g$x_basis)[, which(hal_u_g$coefs != 0)]

  M_U <- cor(as.matrix(x_basis_U))
  corrplot(M_U, method="circle")
}


```


# Linear relationsip & Unpenalized HAL
```{r}
# generate_data_0 <- function(n, a=NA){
#   # exogenous variables
#   U_W <- rnorm(n, 0, 0.1)
#   U_A <- rnorm(n, 0, 0.1)
#   U_Y <- runif(n, 0, 0.1)
#   
#   # endogenous variables
#   W <- U_W
#   
#   if(is.na(a)){
#     A <-  2 - 0.5*W + U_A
#     A[A<=0] = 0
#     A[A>=5] = 5
#   } else {
#     A <- rep(a, n)
#   }
#   
#   Q <- A/5 
#   Q[Q<=0] = 0
#   Q[Q>=1] = 1
#   
#   Y <- rbinom(n, size = 1, prob = Q)
#   
#   # data frame
#   O <- data.frame(W, A, Y)
#   return(O)
# }
# 
# 
# a_vec <- seq(0,5,0.01)
# psi0_a <- c()
# 
# for (i in 1:length(a_vec)) {
#   a <- a_vec[i]
#   
#   EW = 0
#   Q <- a/5 
#   psi0_a[i] = Q
# }

generate_data_0 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 0.1)
  U_A <- rnorm(n, 0, 0.1)
  U_Y <- runif(n, 0, 0.1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- 0.25 * A
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


a_vec <- seq(0,5,0.01)
psi0_a <- c()

for (i in 1:length(a_vec)) {
  a <- a_vec[i]
  
  EW = 0
  psi0_a[i] = a/5
}


psi0_line <- data.frame(a=a_vec, psi0 = psi0_a)

eval_points = seq(0, 5, 0.5)
psi0_pnt <- psi0_line[psi0_line$a %in% eval_points,]

ggplot() +
    geom_line(data=psi0_line, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="a", y="P_0(E[Y|a,W])",
         title = "True Average Treatment Effect, P_0(E[Y|a,W])") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = generate_data_0(n)

hist(obs$Y)
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, # family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, # family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

summary(hal_CV_1)
```
```{r}
summary(hal_CV_1, only_nonzero_coefs = F)
```


```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), # family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), # family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```


# Simu 1 & Unpenalized HAL
```{r}
load(file=here("data", "rdata", paste0("02_simu_V5_sys1_psi0.RData")))

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
source(here("scripts", "scripts_v5_final", "1_hal_functions.R"))

ggplot() +
    geom_line(data=psi0_line, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="a", y="P_0(E[Y|a,W])",
         title = "True Average Treatment Effect, P_0(E[Y|a,W])") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = generate_data_1(n)
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 1)

summary(hal_CV_1)
```

```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```



# Simu 2 & Unpenalized HAL
```{r}
load(file=here("data", "rdata", paste0("02_simu_V5_sys2_psi0.RData")))

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
source(here("scripts", "scripts_v5_final", "1_hal_functions.R"))

ggplot() +
    geom_line(data=psi0_line, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="a", y="P_0(E[Y|a,W])",
         title = "True Average Treatment Effect, P_0(E[Y|a,W])") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = generate_data_2(n)
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 1)

summary(hal_CV_1)
```

```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```

