---
title: "Simulations of estimating causal effects using undersmoothed HAL"
author: "Seraphina Shi"
date: "2023-05-10"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)

library(stringr)
library(glmnet)

library(origami)
library(hal9001)
library(tictoc)

library(R.utils)

library(pROC)

library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)

library(mvtnorm)
```

```{r setup, include = FALSE} 
plotFolder <- here("images")
if(!file.exists(plotFolder)) dir.create(plotFolder,recursive=TRUE)

knitr::opts_chunk$set(
  cache=FALSE, autodep=FALSE, warning=FALSE, message=FALSE, echo=FALSE,
  results = 'markup', dev='png', dpi=150, fig.align = "center", fig.path=paste0(plotFolder, "/"),
  cache.path=".cache/",
  duplicate.label="allow"
)
```


```{r}
R.files.sources <- c(list.files("../R", pattern="*.R$", full.names=TRUE),
                     list.files("../simu_R", pattern="*.R$", full.names=TRUE))
print(R.files.sources)

sapply(R.files.sources, source)
```

```{r}
mean(c(0, 0, 0, 
       0.006, 0.012, 0.018,
       0.020, 0.02, 0.008,
       0.026, 0.030, 0.032))


```


# Simulation 1
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-5 + W + 2.25A -0.5WA)]$

Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys1}
simu.num = 1
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```

# Simulation 2
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1.3^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2.5 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-7 + 3W + 5sin(1.25A^{1.5}) + 5A + 3WA)]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys2}
simu.num = 2
```


```{r child = 'Simu_results_template_one_simu.Rmd'}
```

# Simulation 3
Data structure:  $O = (W, A, Y)$

 * U - exogenous variables
 * W - baseline covariate that is a measure of body condition
 * A - treatment level based on W, continuous between 0 and 5
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-10 - 3W + 4A + \mathbf{I}(A>2) * 5sin((0.8A)^2 - 2.6)  )]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve

```{r check_sys3}
simu.num = 3
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```


# Simulation 4

Data structure:  $O = (W, A, Y)$

 * W - baseline covariates
 * A - treatment level based on W, continuous between 0 and 1
 * Y - outcome, indicator of an event 

 Underlying data generating process, $P_{U,X}$

* Exogenous variables:
  + $U_W \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$
  + $U_Y \sim Uniform(min = 0, max = 1)$

* Structural equations F and endogenous variables:
  + $W =  U_W$
  + $A = bound(2.5 - 0.5W + U_A, min=0, max=5)$
  + $Y = \mathbf{I}[U_Y < expit(-6 + W + 3.5A\mathbf{I}(A\geq2) - 4A\mathbf{I}(A\geq4) - 0.5WA)]$
  
Outcome of interest: $E_0[Y|a,W]$, $a \in [0,5]$, the causal dose-response curve


```{r check_sys4}
simu.num = 4
```

```{r child = 'Simu_results_template_one_simu.Rmd'}
```




# plots and tables 
```{r}
nn = 1000

simu_tbl <- data.frame()

for (a_Val in c(2,4)){
  
  simu.num = 1

  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum) 

  tbl1 = rslt_sum %>% 
  filter(a == a_Val) %>% 
  mutate(simu = 1)
  
  simu.num = 2
  
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum)
  
  tbl2 = rslt_sum %>% 
    filter(a == a_Val) %>% 
  mutate(simu = 2)
  
  simu.num = 3
  
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum)
  
  tbl3 = rslt_sum %>% 
    filter(a == a_Val) %>% 
  mutate(simu = 3)
  
  simu.num = 4
  
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum)
  
  tbl4 = rslt_sum %>% 
    filter(a == a_Val) %>% 
  mutate(simu = 4)
  
  tbl <- rbind(tbl1, tbl2) %>% rbind(tbl3) %>% rbind(tbl4) %>% as.data.frame()
  
  numeric_cols <- tbl %>% select(simu, bias, MSE, cover_rate, bias_se_ratio, SE, oracle_cover_rate, oracle_bias_se_ratio, oracle_SE) %>% round(3)

  # Select non-numeric columns
  non_numeric_cols <- tbl %>% select(method, a) 
  
  # Combine the numeric and non-numeric columns
  sum1_a <- bind_cols(non_numeric_cols, numeric_cols) %>% select(simu, a, method, everything())

  simu_tbl <- rbind(simu_tbl, sum1_a)

}

simu_tbl %>% arrange(simu, a)
```

```{r fig.width=10, fig.height=10}
simu.num = 3
nn = 1000

load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_grid.RData")))

load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
source(here("run_simu", "simu_visual_functions.R"))

rslt_sum = results_1$result_summary
u_g_scaler = mean(unique(rslt_sum$lambda_scaler[rslt_sum$method == 'U_G']))

results_grid_s <- results_grid$result_summary
results_grid_s = add_bound(results_grid_s)


save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_small.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_small.png")))

p <- plot_perforences_grid_small(results_grid_s,
                           u_g_scaler=u_g_scaler, 
                                save_plot= save_plots, 
                                max_bias_sd=NA)


```

```{r, fig.width=15, fig.height=8}
nn = 200

simu.num = 1
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_1 <- results_adapt$result_summary
simu1_plots_200 <- plot_performences_adapt_small(df = results_adapt_1)

simu.num = 3
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_3 <- results_adapt$result_summary
simu3_plots_200 <- plot_performences_adapt_small(df = results_adapt_3)

nn = 5000

simu.num = 1
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_1 <- results_adapt$result_summary
simu1_plots_5000 <- plot_performences_adapt_small(df = results_adapt_1)

simu.num = 3
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_3 <- results_adapt$result_summary
simu3_plots_5000 <- plot_performences_adapt_small(df = results_adapt_3)

for(i in 1:6){
  simu3_plots_200[[i]] <- simu3_plots_200[[i]] + labs(title = " ")
  simu1_plots_5000[[i]] <- simu1_plots_5000[[i]] + labs(title = " ")
  simu3_plots_5000[[i]] <- simu3_plots_5000[[i]] + labs(title = " ")
}

simu_plots = c(simu1_plots_200[1:6], simu1_plots_5000[1:6], simu3_plots_200[1:6], simu3_plots_5000)

g <- grid.arrange(grobs = simu_plots, 
             layout_matrix = rbind(c(1:6),
                                   c(7:12),
                                   c(13:18),
                                   c(19:24),
                                   c(NA,NA,25,25,NA, NA)),
             heights = c(5,5,5,5,1),
             left="            Simulation 3                                                             Simulation 1 \
          n = 5000                            n = 200                             n = 5000                           n = 200 "
             )

save_plot=here(getwd(), "plots", paste0("simu_sys1_sys3_n_200_5000_adapt_performences.png"))
ggsave(save_plot, plot=g, width = 15, height = 8, dpi = 300)

```


```{r, fig.width=11, fig.height=4}
nn = 200

simu.num = 3
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_3 <- results_adapt$result_summary
simu3_plots_200 <- plot_performences_adapt_small(df = results_adapt_3)

nn = 5000

simu.num = 3
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_3 <- results_adapt$result_summary
simu3_plots_5000 <- plot_performences_adapt_small(df = results_adapt_3)

for(i in 1:6){
  simu3_plots_5000[[i]] <- simu3_plots_5000[[i]] + labs(title = " ")
}

simu_plots = c(simu3_plots_200[c(1:3,6)], simu3_plots_5000[c(1:3,6,7)])

g <- grid.arrange(grobs = simu_plots, 
             layout_matrix = rbind(c(1:4),
                                   c(5:8),
                                   c(NA,9,9,NA)),
             heights = c(5,5,1),
             left="                        Simulation 3                 \
                                      n = 5000                            n = 200                             "
             )

save_plot=here(getwd(), "plots", paste0("simu_sys3_n_200_5000_adapt_performences_poster.png"))
ggsave(save_plot, plot=g, width = 11, height = 4, dpi = 300)

```

```{r, fig.width=15, fig.height=8}
nn = 5000
p_list <- list()
results_sum <- data.frame()

check_se_over_1000 <- function(df) {
  if("SE" %in% names(df)) {
    any(df$SE > 1000)
  } else {
    FALSE  # Return FALSE if there is no "SE" column
  }
}

for (simu.num in 1:4) {
    
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_poly.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_GAM.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_npcausal.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  source(here("run_simu", "simu_visual_functions.R"))
  
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum)
  
  results_poly_summary = results_poly$result_summary
  results_poly_summary$method = "Poly"
  results_poly_summary <- rename_df(results_poly_summary)
  
  results_gam_summary = results_gam$result_summary
  results_gam_summary$method = "GAM"
  results_gam_summary <- rename_df(results_gam_summary)
  
  error_ones <- sapply(results_npcausal$all_results, check_se_over_1000)
  results_npcausal$all_results <- results_npcausal$all_results[!error_ones]
  results_npcausal <- run_simu_rep_npcausal_summary(results_npcausal, "npcausal")
  results_npcausal_summary = results_npcausal$result_summary
  results_npcausal_summary <- rename_df(results_npcausal_summary)

  results_df1 = rslt_sum[rslt_sum$method == "U_G",]
  results_df1$method = "HAL"
  
  results_gam_summary$MSE = results_gam_summary$bias^2 + results_gam_summary$oracle_SE^2
  results_poly_summary$MSE = results_poly_summary$bias^2 + results_poly_summary$oracle_SE^2
  results_npcausal_summary$MSE = results_npcausal_summary$bias^2 + results_npcausal_summary$oracle_SE^2
  
  colnames_overlap =intersect(names(results_df1), names(results_npcausal_summary))
  
  results_all <- rbind(results_gam_summary[, colnames_overlap], 
                       results_poly_summary[, colnames_overlap], 
                       results_npcausal_summary[, colnames_overlap], 
                       results_df1[, colnames_overlap])
  
  
  results_all = add_bound(results_all)
  
  p_list[[simu.num]] <- plot_compare_methods_performances_small(results_all)
  
  results_sum_1simu <- results_all %>% 
    mutate(simu = simu.num) %>%
    select(simu, method, bias, MSE, 
           oracle_cover_rate, oracle_bias_se_ratio, oracle_SE, 
           cover_rate, bias_se_ratio, SE) %>%
    group_by(simu, method) %>%
    summarise(across(.cols = where(is.numeric), .fns = mean)) 
    
  
  results_sum <- rbind(results_sum, results_sum_1simu)
  
}

for(i in 1:6){
  p_list[[2]][[i]] <- p_list[[2]][[i]] + labs(title = " ")
  p_list[[3]][[i]] <- p_list[[3]][[i]] + labs(title = " ")
  p_list[[4]][[i]] <- p_list[[4]][[i]] + labs(title = " ")

}

all_plots = c(p_list[[1]][1:6], p_list[[2]][1:6], p_list[[3]][1:6], p_list[[4]])


g <- grid.arrange(grobs = all_plots, 
                   layout_matrix = rbind(c(1:6),
                                   c(7:12),
                                   c(13:18),
                                   c(19:24),
                                   c(NA,NA,25,25,NA, NA)),
             heights = c(5,5,5,5,1), 
             left="             Simulation 4                     Simulation 3                      Simulation 2                    Simulation 1  "
             )

save_plot=here(getwd(), "plots", paste0("simu_sys1234_n", nn,"_compare_performences.png"))
ggsave(save_plot, plot=g, width = 15, height = 8, dpi = 300)


results_sum <- results_sum %>%
  mutate(across(.cols = where(is.numeric), .fns = ~ round(., digits = 3)))

results_sum$method = factor(results_sum$method, levels = c("HAL", "npcausal", "GAM", "Poly"))

results_sum %>% arrange(simu, method)



```



```{r, fig.width=11, fig.height=8}
nn = 5000
p_list <- list()
results_sum <- data.frame()

check_se_over_1000 <- function(df) {
  if("SE" %in% names(df)) {
    any(df$SE > 1000)
  } else {
    FALSE  # Return FALSE if there is no "SE" column
  }
}

for (simu.num in 1:4) {
    
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_poly.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_GAM.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_npcausal.RData")))
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
  source(here("run_simu", "simu_visual_functions.R"))
  
  rslt_sum = results_1$result_summary
  rslt_sum = add_bound(rslt_sum)
  
  results_poly_summary = results_poly$result_summary
  results_poly_summary$method = "Poly"
  results_poly_summary <- rename_df(results_poly_summary)
  
  results_gam_summary = results_gam$result_summary
  results_gam_summary$method = "GAM"
  results_gam_summary <- rename_df(results_gam_summary)
  
  error_ones <- sapply(results_npcausal$all_results, check_se_over_1000)
  results_npcausal$all_results <- results_npcausal$all_results[!error_ones]
  results_npcausal <- run_simu_rep_npcausal_summary(results_npcausal, "npcausal")
  results_npcausal_summary = results_npcausal$result_summary
  results_npcausal_summary <- rename_df(results_npcausal_summary)

  results_df1 = rslt_sum[rslt_sum$method == "U_G",]
  results_df1$method = "HAL"
  
  results_gam_summary$MSE = results_gam_summary$bias^2 + results_gam_summary$oracle_SE^2
  results_poly_summary$MSE = results_poly_summary$bias^2 + results_poly_summary$oracle_SE^2
  results_npcausal_summary$MSE = results_npcausal_summary$bias^2 + results_npcausal_summary$oracle_SE^2
  
  colnames_overlap =intersect(names(results_df1), names(results_npcausal_summary))
  
  results_all <- rbind(results_gam_summary[, colnames_overlap], 
                       results_poly_summary[, colnames_overlap], 
                       results_npcausal_summary[, colnames_overlap], 
                       results_df1[, colnames_overlap])
  
  
  results_all = add_bound(results_all)
  
  p_list[[simu.num]] <- plot_compare_methods_performances_small(results_all)
  
  results_sum_1simu <- results_all %>% 
    mutate(simu = simu.num) %>%
    select(simu, method, bias, MSE, 
           oracle_cover_rate, oracle_bias_se_ratio, oracle_SE, 
           cover_rate, bias_se_ratio, SE) %>%
    group_by(simu, method) %>%
    summarise(across(.cols = where(is.numeric), .fns = mean)) 
    
  
  results_sum <- rbind(results_sum, results_sum_1simu)
  
}

for(i in 1:4){
  p_list[[i]][[1]] <- p_list[[i]][[1]] + 
    labs(title = paste0("Simulation ", i)) +
    theme(plot.title = element_text(hjust = 0.5))
  
  for (j in 2:6) {
    p_list[[i]][[j]] <- p_list[[i]][[j]] + 
      labs(title = " ")
  }

}


all_plots = c(p_list[[1]][c(1:3,6)], p_list[[2]][c(1:3,6)], p_list[[3]][c(1:3,6)], p_list[[4]][c(1:3,6,7)])



g <- grid.arrange(grobs = all_plots, 
                   layout_matrix = rbind(
                                   c(1,5,9,13),
                                   c(2,6,10,14),
                                   c(3,7,11,15),
                                   c(4,8,12,16),
                                   c(NA,17,17,NA)),
             heights = c(5,5,5,5,1)
             )

save_plot=here(getwd(), "plots", paste0("simu_sys1234_n", nn,"_compare_performences_poster.png"))
ggsave(save_plot, plot=g, width = 11, height = 8, dpi = 300)


results_sum <- results_sum %>%
  mutate(across(.cols = where(is.numeric), .fns = ~ round(., digits = 3)))

results_sum$method = factor(results_sum$method, levels = c("HAL", "npcausal", "GAM", "Poly"))

results_sum %>% arrange(simu, method)



```



```{r, fig.width=17, fig.height=5}
simu.num = 1
nn = 200
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_1 <- results_adapt$result_summary
simu1_plots <- plot_performences_adapt_small(df = results_adapt_1)

nn = 1000
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary = add_bound(results_adapt$result_summary)
results_adapt_3 <- results_adapt$result_summary
simu3_plots <- plot_performences_adapt_small(df = results_adapt_3)

for(i in 1:6){
  simu3_plots[[i]] <- simu3_plots[[i]] + labs(title = " ")
}

simu_plots = c(simu1_plots[1:6], simu3_plots)
g <- grid.arrange(grobs = simu_plots, 
             layout_matrix = rbind(c(1,2,3,4,5,6),
                                   c(7,8,9,10,11,12),
                                   c(NA,13,13,13, NA)),
             heights = c(5,5,1),
             left="                Simulation 3                          Simulation 1 \
               "
             )

save_plot=here(getwd(), "plots", paste0("simu_sys1_n_200_1000_adapt_performences.png"))
ggsave(save_plot, plot=g, width = 17, height = 5, dpi = 500)

```


```{r, fig.width=15, fig.height=3}
check_n_basis <- function(df) {
        !all(sapply(df$n_basis, is.numeric) & sapply(df$n_basis, length) == 1)
}

cv_risk_summary_df <- data.frame()
select_1_df <- data.frame()

for (simu.num in c(1,2, 3, 4)) {
  for (nn in c(200, 500, 1000, 5000)) {
    load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
    
    result_list <- results_adapt$result_list
    
    result_summaries <- list()
    methods = names(result_list[[1]])
    
    for (method in c("halcv0_sKnots", "halcv1_sKnots", "halcv2_sKnots", "halcv3_sKnots")){
      result_list_method <- lapply(result_list, function(lst) lst[[method]])
      result_all <-  do.call("rbind", result_list_method) %>% as.data.frame()
      result_all <- merge(as.data.frame(psi0_pnt), result_all, by=c("a"))
      
      data_frames_with_varying_n_basis <- result_list_method[sapply(result_list_method, check_n_basis)]
      
      so <- ifelse(grepl("[0-9]", method), as.numeric(gsub("[^0-9]", "", method)) , -1)
        
      result_summary_cv_risk <- result_all %>% 
        select(cv_risk, if_n_knots_default, smooth_order) %>%
        unique() %>%
        mutate(avg = mean(cv_risk),
               se = sqrt(var(cv_risk))) %>% 
        select(avg, se, if_n_knots_default) %>%
        unique() %>% 
        mutate(ci_lwr = avg - 1.96 * se,
               ci_upr = avg + 1.96 * se,
               so = so,
               simu = simu.num,
               n = nn)
      
      result_summaries[[method]] = result_summary_cv_risk
    }
    
    result_summary = do.call("rbind", result_summaries) %>% 
      as.data.frame() %>%
      select(-if_n_knots_default)
        
    method<- "U_HAL_SL_pick"
  
    cv_risk_summary_df <- rbind(cv_risk_summary_df, result_summary) %>% unique()
      
    
    
    result_list_method <- lapply(result_list, function(lst) lst[[method]])
    result_all_sl <-  do.call("rbind", result_list_method) %>% 
      as.data.frame() %>% 
      filter(a == 0) %>%
      unique() 
    
    select_1 <- mean(result_all_sl$smooth_order == 1)
    
    select_1_df <- rbind(select_1_df, c(simu.num, nn, select_1))
  }
}

cv_risk_summary_df <- cv_risk_summary_df %>% mutate(n = as.factor(n))

p_cv_risk_1 <- ggplot(cv_risk_summary_df %>% filter(simu == 1), 
                    aes(x = so, y = avg, col = n)) +  
      geom_line() + 
      geom_point() + 
      geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, fill=n),  linetype = 0, alpha=0.1) +
      theme_bw()  + 
      labs(x = "smoothness order", y = "CV risk", title = "Simulation 1") 

legend <- get_legend(p_cv_risk_1)
p_cv_risk_1 <- p_cv_risk_1 + theme(legend.position='none')

p_cv_risk_2 <- ggplot(cv_risk_summary_df %>% filter(simu == 2), 
                    aes(x = so, y = avg, col = n)) +  
      geom_line() + 
      geom_point() + 
      geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, fill=n),  linetype = 0, alpha=0.1) +
      theme_bw()  + 
      labs(x = "smoothness order", y = "CV risk", title = "Simulation 2")  + 
  theme(legend.position='none')

p_cv_risk_3 <- ggplot(cv_risk_summary_df %>% filter(simu == 3), 
                    aes(x = so, y = avg, col = n)) +  
      geom_line() + 
      geom_point() + 
      geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, fill=n),  linetype = 0, alpha=0.1) +
      theme_bw()  + 
      labs(x = "smoothness order", y = "CV risk", title = "Simulation 3")  + 
  theme(legend.position='none')

p_cv_risk_4 <- ggplot(cv_risk_summary_df %>% filter(simu == 4), 
                    aes(x = so, y = avg, col = n)) +  
      geom_line() + 
      geom_point() + 
      geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, fill=n),  linetype = 0, alpha=0.1) +
      theme_bw()  + 
      labs(x = "smoothness order", y = "CV risk", title = "Simulation 4")  + 
  theme(legend.position='none')

p <- grid.arrange(p_cv_risk_1, p_cv_risk_2, p_cv_risk_3, p_cv_risk_4, legend, nrow = 1, widths = c(4,4,4,4,1))


ggsave(here(getwd(), "plots", paste0("simu_adapt_cv_risk.png")), plot=p, width = 15, height = 3,  dpi = 300)

```

```{r}
select_1_df %>% pivot_wider(names_from = X200, values_from = X0.706) %>% round(4)
```

```{r}


```

# plots
```{r fig.height=6, fig.width=12, eval=FALSE}
simu.num = 2
load(here("data", "rdata", paste0("simu_", sn, "_", nn, ".RData")))

if(sn %in% c(1,7)){
  load(file=here("data", "rdata", paste0("simu_", sn ,"_psi0.RData")))
  results = run_simu_rep_summary(results)
}

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_2 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_2 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())

legend <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "legend") 




DGS <- generate_data_3
sn = 3
nn=5000
load(here("data", "rdata", paste0("simu_", sn, "_", nn, ".RData")))

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_3 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_3 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())




DGS <- generate_data_7
sn = 7
nn=500
load(here("data", "rdata", paste0("simu_", sn, "_", nn, ".RData")))


if(sn %in% c(1,7)){
  load(file=here("data", "rdata", paste0("simu_", sn ,"_psi0.RData")))
  results = run_simu_rep_summary(results)
}

results$result_summary = add_bound(results$result_summary)
rslt_sum = results$result_summary
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"

source(here("scripts", "scripts_v5_final", "1_visual_functions.R"))
p_est_avg_7 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_est_avg") + theme(axis.title.y=element_blank(), title = element_blank())
p_cr_7 <- plot_performences_cv_ug_alla_noBT(df = rslt_sum, return_plot = "p_cr") + theme(axis.title.y=element_blank(), title = element_blank())



p1 = arrangeGrob(p_est_avg_2, p_est_avg_3, p_est_avg_7, nrow=1, left= "Estimation & CI")
p2 = arrangeGrob(p_cr_2, p_cr_3, p_cr_7, nrow=1, left= "95% CI Coverage Rate")


p <- grid.arrange(p1, p2, legend, 
                  layout_matrix = rbind(c(1,3),
                                        c(2,3)),
                  widths = c(3,0.4),
                  top = "         Simulation 2                                                       Simulation 3                                                        Simulation 4                            \n        n = 500                                                               n = 5000                                                              n = 500                               ")
 

save_plot = here(getwd(), "plots", paste0("simu_sys234_performences_est_cr.png"))
 
ggsave(save_plot, plot=p, width = 12, height = 5, dpi = 800)
```



# colinearity of basis functions
```{r}
library(corrplot)

set.seed(123)
n = 500

U_Y <- runif(n, 0, 1)
U_A <-  rnorm(n, 0, 2)

W <- rnorm(n, 0, 1)

A <-  2 - 0.5*W + U_A
A[A<=0] = 0
A[A>=5] = 5

X <- cbind(W, A)

Y <- as.numeric(U_Y < plogis(-5 + W + 2.25*A - 0.5 * W * A ))
    
obs <- data.frame(W, A, Y)

fit_obj <- fit_hal(X = X, Y = Y, family = "binomial",
                    return_x_basis = TRUE,
                    num_knots = hal9001:::num_knots_generator(
                      max_degree = 1,
                      smoothness_orders = 1
                    ))

summary(fit_obj)

x_basis_CV = cbind(1, hal_CV$x_basis)[, which(hal_CV$coefs != 0)]

M_CV <- cor(as.matrix(x_basis_CV))
corrplot(M_CV, method="circle")
# undersmoothing
```

```{r}
hal_undersmooth <- undersmooth_hal(X, Y, fit_init = hal_CV, family = "binomial")
lambda_u_g = hal_undersmooth$lambda_under

if(!is.na(lambda_u_g)){
  hal_u_g <- fit_hal(X = X, Y = Y, family = "binomial",
                         return_x_basis = TRUE,
                         num_knots = hal9001:::num_knots_generator(
                           max_degree = ifelse(ncol(X) >= 20, 2, 3),
                           smoothness_orders = 1,
                           base_num_knots_0 = 20, #200
                           base_num_knots_1 = 20 # max(100, ceiling(sqrt(n)))
                         ),
                         fit_control = list(
                           cv_select = FALSE,
                           n_folds = 10,
                           foldid = NULL,
                           use_min = TRUE,
                           lambda.min.ratio = 1e-4,
                           prediction_bounds = "default"
                         ),
                         lambda = lambda_u_g)
  
  summary(hal_u_g)
  x_basis_U = cbind(1, hal_u_g$x_basis)[, which(hal_u_g$coefs != 0)]

  M_U <- cor(as.matrix(x_basis_U))
  corrplot(M_U, method="circle")
}
```

```{r}
knitr::knit_exit()
```


# Linear relationsip & Unpenalized HAL
```{r}
# generate_data_0 <- function(n, a=NA){
#   # exogenous variables
#   U_W <- rnorm(n, 0, 0.1)
#   U_A <- rnorm(n, 0, 0.1)
#   U_Y <- runif(n, 0, 0.1)
#   
#   # endogenous variables
#   W <- U_W
#   
#   if(is.na(a)){
#     A <-  2 - 0.5*W + U_A
#     A[A<=0] = 0
#     A[A>=5] = 5
#   } else {
#     A <- rep(a, n)
#   }
#   
#   Q <- A/5 
#   Q[Q<=0] = 0
#   Q[Q>=1] = 1
#   
#   Y <- rbinom(n, size = 1, prob = Q)
#   
#   # data frame
#   O <- data.frame(W, A, Y)
#   return(O)
# }
# 
# 
# a_vec <- seq(0,5,0.01)
# psi0_a <- c()
# 
# for (i in 1:length(a_vec)) {
#   a <- a_vec[i]
#   
#   EW = 0
#   Q <- a/5 
#   psi0_a[i] = Q
# }

generate_data_0 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 0.1)
  U_A <- rnorm(n, 0, 0.1)
  U_Y <- runif(n, 0, 0.1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- 0.25 * A
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}


a_vec <- seq(0,5,0.01)
psi0_a <- c()

for (i in 1:length(a_vec)) {
  a <- a_vec[i]
  
  EW = 0
  psi0_a[i] = a/5
}


psi0_line <- data.frame(a=a_vec, psi0 = psi0_a)

eval_points = seq(0, 5, 0.5)
psi0_pnt <- psi0_line[psi0_line$a %in% eval_points,]

ggplot() +
    geom_line(data=psi0_line, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="a", y="P_0(E[Y|a,W])",
         title = "True Average Treatment Effect, P_0(E[Y|a,W])") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = generate_data_0(n)

hist(obs$Y)
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, # family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, # family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

summary(hal_CV_1)
```
```{r}
summary(hal_CV_1, only_nonzero_coefs = F)
```


```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), # family = "binomial",
                                return_x_basis = TRUE,
                                smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), # family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```


# Simu 1 & Unpenalized HAL
```{r}
load(file=here("data", "rdata", paste0("simu_1_psi0.RData")))

R.files.sources <- c(list.files("../R", pattern="*.R$", full.names=TRUE),
                     list.files("../simu_R", pattern="*.R$", full.names=TRUE))
print(R.files.sources)

sapply(R.files.sources, source)

ggplot() +
    geom_line(data=psi0_all, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="Treatment", y="Outcome",
         title = "True Causal Dose Response Curve") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = DGS(simu=1, n) 
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 1)

summary(hal_CV_1)
```

```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```



# Simu 2 & Unpenalized HAL
```{r}
load(file=here("data", "rdata", paste0("simu_2_psi0.RData")))

R.files.sources <- c(list.files("../R", pattern="*.R$", full.names=TRUE),
                     list.files("../simu_R", pattern="*.R$", full.names=TRUE))
print(R.files.sources)

sapply(R.files.sources, source)

ggplot() +
    geom_line(data=psi0_all, aes(x=a, y=psi0)) + 
    geom_point(data=psi0_pnt, aes(x=a, y=psi0)) + 
    labs(x="Treatment", y="Outcome",
         title = "True Causal Dose-Response Curve") +
        theme(plot.title = element_text(hjust = 0.5),
              panel.grid.major.x = element_blank(),
              axis.text = element_text(size=7)) +
  theme_bw()
```


```{r}
set.seed(123)

n = 200
obs = DGS(simu=2, n)
```

```{r}
y_name = "Y"
x_names = names(obs)[names(obs) != 'Y']
  
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
```

```{r}
hal_CV_0 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 0)

summary(hal_CV_0)
```

```{r}
pred_rslts_0 = predict_curve(X, Y, hal_fit=hal_CV_0, points_curve = eval_points)
```



```{r}
hal_CV_1 <- fit_hal(X = X, Y = Y, family = "binomial",
                  return_x_basis = TRUE,
                  smoothness_orders = 1)

summary(hal_CV_1)
```

```{r}
pred_rslts_1 = predict_curve(X, Y, hal_fit=hal_CV_1, points_curve = eval_points)
```


## Add unpenalized X
```{r}
hal_CV_unpenalized_0 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 0)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_0)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_0 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_0, points_curve = eval_points)
```

```{r}
hal_CV_unpenalized_1 <- fit_hal(Y = Y, X = X, X_unpenalized = as.matrix(X), family = "binomial",
                    return_x_basis = TRUE,
                    smoothness_orders = 1)

sum <- summary_hal9001_unpenalized(hal_CV_unpenalized_1)
print(sum$table)
print(sum$table[order(abs(sum$table$coef), decreasing = T), ])
```

```{r}
pred_rslts_unpenalized_1 = predict_curve(X, Y, hal_fit=hal_CV_unpenalized_1, points_curve = eval_points)
```

## plot results
```{r}
pred_rslts_0$method = "0_hal"
pred_rslts_1$method = "1_hal"
pred_rslts_unpenalized_0$method = "0_hal_unpenalized"
pred_rslts_unpenalized_1$method = "1_hal_unpenalized"

pred_table = rbind(pred_rslts_0, pred_rslts_1) %>% rbind(pred_rslts_unpenalized_0) %>% rbind(pred_rslts_unpenalized_1) %>% merge(psi0_pnt, by = "a")
print(pred_table)

```

```{r}
p_est_avg <- ggplot(data=pred_table, aes(x=a)) +
    geom_line(aes(y=psi0), alpha = 0.5, color="darkgrey") +
    geom_ribbon(aes(ymin=ci_lwr, ymax=ci_upr, color=method, fill=method), width=0.7, alpha=0.1) +
    geom_point(aes(y=psi0), color = "black") +
    geom_point(aes(y=y_hat, color=method), shape=17, size=2, alpha= 0.7) +
    labs(x="a", y="E[Y|a, W]", title = "Estimation") +
    theme_bw() +
    theme(legend.box = "horizontal")
p_est_avg
```

