
### CV vs Undersmoothing (Zero smoothness order with default number of knots)

```{r}
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_zero.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_0$result_summary <- rename_df(results_0$result_summary)

rslt_sum_zero = results_0$result_summary

cv_lambda = unique(rslt_sum_zero$lambda[rslt_sum_zero$method == 'CV'])
u_g_lambda = unique(rslt_sum_zero$lambda[rslt_sum_zero$method == 'U_G'])
u_g_scaler = unique(rslt_sum_zero$lambda_scaler[rslt_sum_zero$method == 'U_G'])

eval_points = seq(0,5,0.5)

cv_n_bases = unique(rslt_sum_zero$n_basis[rslt_sum_zero$method == 'CV'])
u_g_n_bases = unique(rslt_sum_zero$n_basis[rslt_sum_zero$method == 'U_G'])

cat(sprintf(' The average lambda of CV-HAL: %0.4f (= 1 * lambda_CV )', cv_lambda))
cat(sprintf(' The average lambda of globally undersmoothed HAL: %0.4f (= %0.4f * lambda_CV )', u_g_lambda, u_g_scaler))

cat(sprintf(' The average number of bases of CV-HAL: %0.4f ', cv_n_bases))
cat(sprintf(' The average number of bases of globally undersmoothed HAL: %0.4f ', u_g_n_bases))


cat(sprintf(' The average fitting time for CV-HAL: %0.4f seconds', mean(rslt_sum_zero$hal_fit_time[rslt_sum_zero$method=='CV'])))
cat(sprintf(' The average fitting time for globally undersmoothed HAL: %0.4f seconds', mean(rslt_sum_zero$hal_fit_time[rslt_sum_zero$method=='U_G'])))


```

```{r fig.height=5, fig.width=10}
rslt_sum_zero$method[rslt_sum_zero$method == "U_G"] = "Undersmooth"

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "zero_HAL", paste0("simu_sys", simu.num, "_n", n, "_performences_zero.png")),
                here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_performences_zero.png")))

p <- plot_performences_cv_ug_alla(df = rslt_sum_zero, save_plot = save_plots)
```

```{r fig.height=4, fig.width=12}
save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "zero_HAL", paste0("simu_sys", simu.num, "_n", n, "_QQplots_zero.png")),
                here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_QQplots_zero.png"))
                )

p <- estimation_qqplot_cv_ug_alla(results_list = results_0, save_plot = save_plots)
plot(p)
```


```{r fig.height=4, fig.width=15}
p_est_avg <- plot_performences_cv_ug_alla(df = rslt_sum_zero, return_plot = "p_est_avg") + 
  theme(axis.title.y=element_blank(), plot.title = element_text(hjust = 0.5)) + 
  labs(title = "(a) Estimations & 95% CIs")
p_cr <- plot_performences_cv_ug_alla(df = rslt_sum_zero, return_plot = "p_cr") + 
  theme(axis.title.y=element_blank(), plot.title = element_text(hjust = 0.5)) + 
  labs(title = "(b) 95% CI coverage rates")
p_bias_se <- plot_performences_cv_ug_alla(df = rslt_sum_zero, return_plot = "p_bias_se") + 
  theme(axis.title.y=element_blank(), plot.title = element_text(hjust = 0.5)) + 
  labs(title = "(c) |Bias| / Standard Error")

legend <- plot_performences_cv_ug_alla(df = rslt_sum_zero, return_plot = "legend") 

p <- grid.arrange(p_est_avg, p_cr, p_bias_se, legend,
                  layout_matrix = matrix(c(1,2,3,4), nrow=1),  widths = c(1,1,1,0.4))

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "zero_HAL", paste0("simu_sys", simu.num, "_n", n, "_performences_zero_small.png")),
                here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_performences_zero_small.png"))
                )
for (i in 1:length(save_plots)) {
  save_loc = save_plots[i]
  ggsave(save_loc, plot=p, width = 12, height = 3, dpi = 800)
}

  
```

### CV vs Undersmoothing (First smoothness order with smaller number of knots)

```{r}
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first.RData")))
source(here("run_simu", "simu_visual_functions.R"))

rslt_sum = results_1$result_summary

cv_lambda = unique(rslt_sum$lambda[rslt_sum$method == 'CV'])
u_g_lambda = mean(unique(rslt_sum$lambda[rslt_sum$method == 'U_G']))
u_g_scaler = mean(unique(rslt_sum$lambda_scaler[rslt_sum$method == 'U_G']))

eval_points = seq(0,5,0.5)

cv_n_bases = unique(rslt_sum$n_basis[rslt_sum$method == 'CV'])
u_g_n_bases = unique(rslt_sum$n_basis[rslt_sum$method == 'U_G'])

cat(sprintf(' The average lambda of CV-HAL: %0.4f (= 1 * lambda_CV )', cv_lambda))
cat(sprintf(' The average lambda of globally undersmoothed HAL: %0.4f (= %0.4f * lambda_CV )', u_g_lambda, u_g_scaler))

cat(sprintf(' The average number of bases of CV-HAL: %0.4f ', cv_n_bases))
cat(sprintf(' The average number of bases of globally undersmoothed HAL: %0.4f ', u_g_n_bases))


cat(sprintf(' The average fitting time for CV-HAL: %0.4f seconds', mean(rslt_sum$hal_fit_time[rslt_sum$method=='CV'])))
cat(sprintf(' The average fitting time for globally undersmoothed HAL: %0.4f seconds', mean(rslt_sum$hal_fit_time[rslt_sum$method=='U_G'])))


```

```{r fig.height=5, fig.width=10}
rslt_sum$method[rslt_sum$method == "U_G"] = "Undersmooth"


save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "first_HAL", paste0("simu_sys", simu.num, "_n", n, "_performences.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_performences.png"))
                )
p <- plot_performences_cv_ug_alla(df = rslt_sum, save_plot =save_plots)
```

```{r fig.height=4, fig.width=12}

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "first_HAL", paste0("simu_sys", simu.num, "_n", n, "_QQplots.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_QQplots.png")))
                
p <- estimation_qqplot_cv_ug_alla(results_list = results_1, save_plot = save_plots)
plot(p)
```


### CV vs Undersmoothing (First smoothness order with even smaller number of knots)

```{r}
if (nn != 5000){
  load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_first_smallKnot.RData")))
  source(here("run_simu", "simu_visual_functions.R"))
  
  rslt_smallKnots_sum = results_1_smallKnots$result_summary
  
  cv_lambda_smallKnots = unique(rslt_smallKnots_sum$lambda[rslt_smallKnots_sum$method == 'CV'])
  u_g_lambda_smallKnots = mean(unique(rslt_smallKnots_sum$lambda[rslt_smallKnots_sum$method == 'U_G']))
  u_g_scaler_smallKnots = mean(unique(rslt_smallKnots_sum$lambda_scaler[rslt_smallKnots_sum$method == 'U_G']))
  
  eval_points = seq(0,5,0.5)
  
  cv_n_bases_smallKnots = unique(rslt_smallKnots_sum$n_basis[rslt_smallKnots_sum$method == 'CV'])
  u_g_n_bases_smallKnots = unique(rslt_smallKnots_sum$n_basis[rslt_smallKnots_sum$method == 'U_G'])
  
  cat(sprintf(' The average lambda of CV-HAL: %0.4f (= 1 * lambda_CV )', cv_lambda_smallKnots))
  cat(sprintf(' The average lambda of globally undersmoothed HAL: %0.4f (= %0.4f * lambda_CV )', u_g_lambda_smallKnots, u_g_scaler_smallKnots))
  
  cat(sprintf(' The average number of bases of CV-HAL: %0.4f ', cv_n_bases_smallKnots))
  cat(sprintf(' The average number of bases of globally undersmoothed HAL: %0.4f ', u_g_n_bases_smallKnots))
  
  
  cat(sprintf(' The average fitting time for CV-HAL: %0.4f seconds', mean(rslt_smallKnots_sum$hal_fit_time[rslt_smallKnots_sum$method=='CV'])))
  cat(sprintf(' The average fitting time for globally undersmoothed HAL: %0.4f seconds', mean(rslt_smallKnots_sum$hal_fit_time[rslt_smallKnots_sum$method=='U_G'])))
}



```

```{r fig.height=5, fig.width=10}
if (nn != 5000){
  rslt_smallKnots_sum$method[rslt_smallKnots_sum$method == "U_G"] = "Undersmooth"


  save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "first_HAL", paste0("simu_sys", simu.num, "_n", n, "_performences_smallKnots.png")), 
                   here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_performences_smallKnots.png"))
                  )
  p <- plot_performences_cv_ug_alla(df = rslt_smallKnots_sum, save_plot =save_plots)
}

```

```{r fig.height=4, fig.width=12}
if (nn != 5000){
  save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "first_HAL", paste0("simu_sys", simu.num, "_n", n, "_QQplots_smallKnots.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_QQplots_smallKnots.png")))
                
p <- estimation_qqplot_cv_ug_alla(results_list = results_1_smallKnots, save_plot = save_plots)
plot(p)
}

```

### small knots vs smaller knots (First smoothness order)
```{r fig.height=5, fig.width=10}
if (nn != 5000){
  rslt_sum_tmp <- rslt_sum
rslt_sum_tmp$method <- paste0(rslt_sum_tmp$method, "_knots20")

rslt_smallKnots_sum_tmp <- rslt_smallKnots_sum
rslt_smallKnots_sum_tmp$method <- paste0(rslt_smallKnots_sum_tmp$method, "_knots10")

rslt_sum_all <- rbind(rslt_sum_tmp, rslt_smallKnots_sum_tmp)


save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "first_HAL", paste0("simu_sys", simu.num, "_n", n, "_performences_compareKnots.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_performences_compareKnots.png"))
                )
p <- plot_performences_knots(df = rslt_sum_all, save_plot =save_plots)
}

```


### Undersmoothed Smoothness-adaptive HAL
```{r fig.width=7, fig.height=14}
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_adapt.RData")))
source(here("run_simu", "simu_visual_functions.R"))
results_adapt$result_summary <- rename_df(results_adapt$result_summary)

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "U_SOadaptive_HAL", paste0("simu_sys", simu.num, "_n", n, "_adapt_performences.png")),
                save_plot=here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_adapt_performences.png"))
                )

p <- plot_performences_adapt(df = results_adapt$result_summary,
                             save_plots)

```

```{r}
df_adapt_n_basis = results_adapt$result_summary %>% select(smooth_order, if_n_knots_default, sl_pick, lambda, lambda_scaler, n_basis, cv_risk) %>% unique()

df_adapt_n_basis = round(df_adapt_n_basis, 4)
df_adapt_n_basis$if_n_knots_default = factor(df_adapt_n_basis$if_n_knots_default)
print(df_adapt_n_basis)


df_adapt_n_basis$label_val = as.character(round(df_adapt_n_basis$n_basis, 2)) 
df_adapt_n_basis$lambda_val = as.character(round(df_adapt_n_basis$lambda, 2)) 
df_adapt_n_basis$cv_risk_val = as.character(round(df_adapt_n_basis$cv_risk, 2)) 


p_n_basis_grid <- ggplot(df_adapt_n_basis, aes(x = smooth_order, color = if_n_knots_default)) +  
      geom_line(aes(y = n_basis, color = if_n_knots_default)) + 
      geom_point(aes(y = n_basis, color = if_n_knots_default)) + 
  geom_text(aes(y = n_basis, label = label_val), hjust=0, vjust=0) + 
      theme_bw()  + 
      labs(x = "smooth order", y = "Number of bases") 
p_n_basis_grid


ggsave(here(getwd(), "plots", paste0("sys", simu.num), "U_SOadaptive_HAL", paste0("simu_sys", simu.num, "_n", n, "_adapt_num_basis.png")), plot=p_n_basis_grid, width = 5, height = 3,  dpi = 800)
ggsave(here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_adapt_num_basis.png")), plot=p_n_basis_grid, width = 5, height = 3,  dpi = 800)

p_lambda_grid <- ggplot(df_adapt_n_basis, aes(x = smooth_order, color = if_n_knots_default)) +  
      geom_line(aes(y = lambda, color = if_n_knots_default)) + 
      geom_point(aes(y = lambda, color = if_n_knots_default)) + 
  geom_text(aes(y = lambda, label = lambda_val), hjust=0, vjust=0) + 
      theme_bw()  + 
      labs(x = "smooth order", y = "lambda") 
p_lambda_grid
ggsave(here(getwd(), "plots", paste0("sys", simu.num), "U_SOadaptive_HAL", paste0("simu_sys", simu.num, "_n", n, "_adapt_lambda.png")), plot=p_lambda_grid, width = 5, height = 3,  dpi = 800)
ggsave(here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_adapt_lambda.png")), plot=p_lambda_grid, width = 5, height = 3,  dpi = 800)

p_cv_risk_grid <- ggplot(df_adapt_n_basis, aes(x = smooth_order, color = if_n_knots_default)) +  
      geom_line(aes(y = cv_risk, color = if_n_knots_default)) + 
      geom_point(aes(y = cv_risk, color = if_n_knots_default)) + 
  geom_text(aes(y = cv_risk, label = cv_risk_val), hjust=0, vjust=0) + 
      theme_bw()  + 
      labs(x = "smooth order", y = "CV risk") 
p_cv_risk_grid
ggsave(here(getwd(), "plots",paste0("sys", simu.num), "U_SOadaptive_HAL",  paste0("simu_sys", simu.num, "_n", n, "_adapt_cv_risk.png")), plot=p_cv_risk_grid, width = 5, height = 3,  dpi = 800)
ggsave(here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_adapt_cv_risk.png")), plot=p_cv_risk_grid, width = 5, height = 3,  dpi = 800)

```


### grid of scalers
```{r fig.width=25, fig.height=9}
# lambda_scalers = c(1.2, 1.1, 10^seq(from=0, to=-3, length=20))

load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_grid.RData")))

source(here("run_simu", "simu_visual_functions.R"))

# results_grid_s <- results_grid_summary(results_grid)

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences.png")))

p <- plot_perforences_grid(results_grid_s,
                           u_g_scaler=u_g_scaler, 
                                save_plot= save_plots, 
                                max_bias_sd=NA)

no_empirical_CI_proportion <- results_grid$no_empirical_CI_proportion
lambda_scalers <- unique(results_grid_s$lambda_scaler)

names(no_empirical_CI_proportion) = as.character(round(lambda_scalers,4))
print("proportion of simulations that cannot compute empirical SD:")
print(no_empirical_CI_proportion)
```

```{r}
df_grid_n_basis = unique(results_grid_s[, c('lambda_scaler', 'n_basis')])

cv_n_b = mean(df_grid_n_basis$n_basis[df_grid_n_basis$lambda_scaler == 1])
u_g_n_b = mean(df_grid_n_basis$n_basis[which(abs(df_grid_n_basis$lambda_scaler - u_g_scaler) == min(abs(df_grid_n_basis$lambda_scaler - u_g_scaler) ))])

p_n_basis_grid <- ggplot(df_grid_n_basis, aes(x = lambda_scaler)) +  
      geom_line(aes(y = n_basis)) + 
      geom_point(aes(y = n_basis)) + 
      theme_bw()  + 
      labs(x = "lambda scalers", y = "Number of bases") +
      theme(legend.position='none') +
        geom_vline(xintercept = u_g_scaler, lty=2, col = "deepskyblue") +
        geom_vline(xintercept = 1, lty=2, col = "purple") + 
     annotate(x=u_g_scaler, y=+Inf, label=round(u_g_n_b, 2), vjust=2, geom="label") +
     annotate(x=1, y=+Inf, label=round(cv_n_b, 2), vjust=2, geom="label")


p_n_basis_grid

ggsave(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers",paste0("simu_sys", simu.num, "_n", n, "_grid_num_basis.png")), plot=p_n_basis_grid, width = 5, height = 3,  dpi = 800)
ggsave(here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_num_basis.png")), plot=p_n_basis_grid, width = 5, height = 3,  dpi = 800)


```

### grid of extra scalers
```{r fig.width=25, fig.height=9}
# lambda_scalers = c(1.2, 1.1, 10^seq(from=0, to=-3, length=20))

if (nn != 5000 & simu.num != 4){
  
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_grid_extra.RData")))
source(here("run_simu", "simu_visual_functions.R"))

# results_grid_all <- results_grid_summary(results_grid_extra)
# results_grid_all <- rename_df(results_grid_all)
# results_grid_all = add_bound(results_grid_all)
results_grid_all$method[results_grid_all$method == "U_G"] = "Undersmooth"

# results_grid_all$MSE = results_grid_all$bias^2 + results_grid_all$SE^2
# results_grid_all$oracle_MSE = results_grid_all$bias^2 + results_grid_all$oracle_SE^2
results_grid_all$lambda = results_grid_all$lambda_scaler * mean(results_grid_all$lambda[results_grid_all$lambda_scaler == 1])

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra.png")))

p <- plot_perforences_grid(results_grid_all,
                           u_g_scaler=u_g_scaler, 
                                save_plot=save_plots, 
                                max_bias_sd=NA)
}

```


```{r fig.width=25, fig.height=11}
if (nn != 5000 & simu.num != 4){
  save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_lambda.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_lambda.png")))


p <- plot_perforences_grid_lambda(results_grid_all,
                                                u_g_lambda = u_g_lambda, 
                                                  cv_lambda = cv_lambda,
                                                  save_plot=save_plots,
                                                  max_bias_sd=NA )
}

```

```{r fig.height=5, fig.width=10}
if (nn != 5000 & simu.num != 4){
  save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_a1.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_a1.png")))

p <- plot_perforences_grid_lambda_a(results_grid_all,
                                                  a = 1,
                                                  u_g_lambda = u_g_lambda, 
                                                  cv_lambda = cv_lambda,
                                                  save_plot=save_plots,
                                                  max_bias_sd=NA )

save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "grid_scalers", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_a2.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_grid_performences_extra_a2.png")))
p <- plot_perforences_grid_lambda_a(results_grid_all,
                                                  a = 2,
                                                  u_g_lambda = u_g_lambda, 
                                                  cv_lambda = cv_lambda,
                                                  save_plot=save_plots,
                                                  max_bias_sd=NA )
}

```


### vs GAM & Polynomial regression
```{r}
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_poly.RData")))
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_GAM.RData")))
load(here("data", "rdata", paste0("simu_", simu.num, "_", nn, "_npcausal.RData")))


results_npcausal <- run_simu_rep_npcausal_summary(results_npcausal, "npcausal")

results_poly$result_summary$method = "Poly"

results_df1 = rslt_sum_zero[rslt_sum_zero$method == "CV",]
results_df1$method = "0_HAL"
results_df2 = results_adapt$result_summary[results_adapt$result_summary$method == "U_HAL_SL_pick", ]
results_df2$method = "U_SOadapt_HAL"

results_gam$result_summary <- rename_df(results_gam$result_summary)
results_poly$result_summary <- rename_df(results_poly$result_summary)
results_npcausal$result_summary <- rename_df(results_npcausal$result_summary)

colnames_overlap = intersect( intersect(names(results_df1), names(results_df2)), names(results_poly$result_summary))

results_all <- rbind(results_gam$result_summary[, colnames_overlap], 
                     results_poly$result_summary[, colnames_overlap], 
                     results_npcausal$result_summary[, colnames_overlap], 
                     results_df1[, colnames_overlap],
                     results_df2[, colnames_overlap])


results_all = add_bound(results_all)

results_all$MSE = results_all$bias^2 + results_all$SE^2
results_all$oracle_MSE = results_all$bias^2 + results_all$oracle_SE^2
```


```{r fig.height=15, fig.width=8}
save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "compare_methods", paste0("simu_sys", simu.num, "_n", n, "_compare_performances.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_compare_performances.png")))

p <- plot_compare_methods_performances(results_all,save_plot = save_plots )
```

```{r fig.height=15, fig.width=8}
save_plots <- c(here(getwd(), "plots", paste0("sys", simu.num), "compare_methods", paste0("simu_sys", simu.num, "_n", n, "_compare_performances_npcausal.png")), 
                 here(getwd(), "plots", paste0("simu_sys", simu.num, "_n", n, "_compare_performances_npcausal.png")))


p <- plot_compare_methods_performances_npcausal(results_all,save_plot = save_plots )
```



